Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massive Distributed Evolutionary Computation

### 2015-08-08 volunteer-clouds-ieee-ci-si.tex
Revert "More about the algorithm"
This reverts commit 4e35e721982c6ae9b9bccaf38301da7457f4a2cb.

JavaScript has commonly been used to develop client-side scripts that handle the user interface,
communicate asynchronously with servers and update the content that is displayed \cite{flanagan2006javascript}.
Long-running scripts like those proposed in this work can interfere with the responsiveness of 
the web application, because everything is handled by a single thread. 
The Web Workers specification \cite{hickson2012web} defines an API that allows Web developers 
to execute long-running scripts in parallel with the scripts of the main page. 
Worker scripts are spawned in the background allowing a thread-like operation using 
message-passing as the communication mechanism between the worker's execution environment and the main thread. 
According to the Web Workers W3C Candidate Recomendation \cite{hickson2012web} 
workers are expected to be long-lived, they have a high start-up performance cost, 
and a high per-instance memory cost. The Web Worker API is implemented in current
versions of both desktop and mobile web browsers. For volunteer computing using the Web Worker API   
brings several advantages over a single threaded implementation:
\begin{itemize}
\item If the browser uses a tabbed document interface, the worker script 
keeps running in the background even it the user brings forward another tabbed window.
\item Several instances of the NodEO script can be executed in parallel in a single web page. 
\item Many immplemantations of the Web Worker API can use multiple CPUs for the execution
of worker scripts, in particular the execution in multiple CPUs was tested in desktop verions of
Chrome, FireFox and Safari.

\end{itemize}

\begin{figure}[!t]
\centering
\includegraphics[width=4in]{img/Algorithm.pdf}
\caption{Description of the W$^2$ version of {\sf NodIO}. In this
  version, clientts use WebWorkers to run the evolutionary algorithm,
  with two of them per browser.}  
\label{fig:system:w2}
\end{figure}

>>>>

JavaScript has commonly been used to develop client-side scripts that handle the user interface,
communicate asynchronously with servers and update the content that is displayed \cite{flanagan2006javascript}.
Long-running scripts like those proposed in this work can interfere with the responsiveness of 
the web application, because everything is handled by a single thread. 
The Web Workers specification \cite{hickson2012web} defines an API that allows Web developers 
to execute long-running scripts in parallel with the scripts of the main page. 
Worker scripts are spawned in the background allowing a thread-like operation using 
message-passing as the communication mechanism between the worker's execution environment and the main thread. 
According to the Web Workers W3C Candidate Recomendation \cite{hickson2012web} 
workers are expected to be long-lived, they have a high start-up performance cost, 
and a high per-instance memory cost. The Web Worker API is implemented in current
versions of both desktop and mobile web browsers. For volunteer computing using the Web Worker API   
brings several advantages over a single threaded implementation:

\begin{itemize}
\item If the browser uses a tabbed document interface, the worker script 
keeps running in the background even it the user brings forward another tabbed window.
\item Several instances of NodEO can be executed in parallel in a single web page. 
\item Many immplemantations of the Web Worker API can use multiple CPUs for the execution
of worker scripts, in particular the execution in multiple CPUs was tested in desktop verions of
Chrome, FireFox and Safari.
\item Each NodEO instance can be restarted independenly.
\end{itemize}

A sequence diagram of the W$^2$ version of {\sf NodIO} is presented in Figure~\ref{fig:system:w2}. In this
version, clients use the Web Workers API to run two NodEO instances per browser. The sequence diagram shows
the interaction between processess in a time sequence. In this system there are two kinds of asynchronus
messages: HTTP Requests shown as bold arrows between the client and the server where PUT and GET HTTP Methods
are represented by red and blue arrows respectively. There are also messages between workers and the main thread in
locally in the client, this are shown as thin solid arrows. There are three processes involved:

\begin{itemize}
\item {\em nodejs} This is the server side process, implented in the nodejs framework. Web request handlers are
implemented using the Express web framework. For keeping a log of the execution the Winston library is used, and finally
the NodeIO framework is responsible for keeping the shared pool.
\item {\em Client main thread}  This process handles the user interaction using the JQuery framwork. The Chart.js library
is used to display a chart of the best fitness in the population over time. The main script also creates 
worker instances and handles the user interface.  
\item {\em Worker global scope} The worker execution space does not have access to the Document Object Model (DOM) 
or JQuery objects in the main thread. Asynchronus HTTP requests to the server are implemented using the XHR object. 
\end{itemize}

The sequence of events are:
\begin{enumerate}
\item First the NodIO process is started in nodejs. As part of the initialization procedure the number
of experiment is initialised and a log file is created. The shared pool implemented as an array is reset.
\item A volunteer follows the link of the experiment triggering a GET HTTP request and the page is 
loaded.
\item The main script renders the page and creates the worker 
\end{enumerate}

\begin{figure}[!t]
\centering
\includegraphics[width=4in]{img/Algorithm.pdf}
\caption{Description of the W$^2$ version of {\sf NodIO}. In this
  version, clients use Web Workers to run the evolutionary algorithm,
  with two of them per browser.}  
\label{fig:system:w2}
\end{figure}
